---
title: "Compiling Summary Statistics for the Cook County Medical Respite Center"
author: "Daniel P. Hall Riggins, MD"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

# Load needed libraries:

```{r message=FALSE, warning=FALSE}
library(arrow)
library(lubridate)
library(tidyverse)
```

```{r}
Sys.unsetenv("RETICULATE_PYTHON")
library(reticulate)
#install_miniconda()
#conda_create("r_mrc_analyses")
use_miniconda(condaenv = "r_mrc_analyses", required = TRUE)
py_config()
#py_install(packages = "pandas-dedupe", pip = TRUE)
```
```{python}
import pandas as pd
import pandas_dedupe as dd
import os
```

# Import data from RedCap:

```{r}
all_raw <- read_csv("data/MRC_ALL_DATA_LABELS_2021-09-30_1556.csv")
```

Calculate total months of operations:

```{r}
start_date <- as_date("2020-12-08") # From date of first communication
date_of_export <- as_date("2021-09-30") # From time-stamp on data file
operation_months <- start_date %--% date_of_export / dmonths(1)
print(
    paste0(
        "This data download represents ",
        floor(operation_months),
        " total months of operation."
    )
)
```

To get total referrals (174), get the max number for client ID:

```{r}
print(
    paste0(
        "There have been ",
        max(all_raw$`Record ID`),
        " total referrals."
    )
)
```

Collect referral-level data:

```{r}
referrals <- all_raw %>%
    select(
        record = `Record ID`, 
        referral_source = `Referring Organization`,
        primary_ref_type = `Primary Referral Category`,
        disposition = `Disposition?` # Aka whether referral was accepted
    ) %>%
    # Only keep records that have data present about one of the referral variables
    filter(
        !is.na(referral_source) | !is.na(primary_ref_type) | !is.na(disposition)
    ) %>%
    # Coerce variables into un-ordered factors 
    mutate(
        referral_source = str_to_lower(referral_source),
        # Consolidate referrals source categories using string matching
        referral_source_cat = as_factor(
            case_when(
                str_detect(
                    string = referral_source,
                    pattern = "macneal"
                ) ~ "MacNeal Hospital, Loyola Medicine",
                str_detect(
                    string = referral_source,
                    pattern = "loyola"
                ) ~ "Loyola UMC",
                str_detect(
                    string = referral_source,
                    pattern = "stroger"
                ) ~ "John H. Stroger Hospital, Cook County Health",
                str_detect(
                    string = referral_source,
                    pattern = "osf"
                ) ~ "OSF Little Company of Mary Medical Center",
                str_detect(
                    string = referral_source,
                    pattern = "olympia"
                ) ~ "Franciscan Health Olympia Fields",
                str_detect(
                    string = referral_source,
                    pattern = "alsip"
                ) ~ "CCDPH Alsip Alternate Housing Program",
                str_detect(
                    string = referral_source,
                    pattern = "ridge"
                ) ~ "Advocate Lutheran General Hospital",
                str_detect(
                    string = referral_source,
                    pattern = "swedish"
                ) ~ "Swedish Hospital, NorthShore",
                str_detect(
                    string = referral_source,
                    pattern = "gottlieb"
                ) ~ "Gottlieb Memorial Hospital, Loyola Medicine",
                str_detect(
                    string = referral_source,
                    pattern = "palos"
                ) ~ "Palos Health, Northwestern Medicine",
                str_detect(
                    string = referral_source,
                    pattern = "oak"
                ) ~ "Rush Oak Park Hospital",
                str_detect(
                    string = referral_source,
                    pattern = "connections"
                ) ~ "Connections for the Homeless",
                str_detect(
                    string = referral_source,
                    pattern = "provident"
                ) ~ "Provident Hospital of Cook County",
                str_detect(
                    string = referral_source,
                    pattern = "amita"
                ) ~ "AMITA Health Saint Francis Hospital Evanston",
                str_detect(
                    string = referral_source,
                    pattern = "rush u"
                ) ~ "Rush University Medical Center",
                str_detect(
                    string = referral_source,
                    pattern = "respond"
                ) ~ "Respond Now",
                str_detect(
                    string = referral_source,
                    pattern = "gateway"
                ) ~ "Gateway Foundation",
                str_detect(
                    string = referral_source,
                    pattern = "northwestern memorial"
                ) ~ "Northwestern Memorial Hospital",
                str_detect(
                    string = referral_source,
                    pattern = "safe"
                ) ~ "A Safe Haven",
                str_detect(
                    string = referral_source,
                    pattern = "defend"
                ) ~ "Law Office of the Cook COunty Public Defender",
                str_detect(
                    string = referral_source,
                    pattern = "access"
                ) ~ "Access Community Health Network",
                str_detect(
                    string = referral_source,
                    pattern = "sojou"
                ) ~ "Sojourner House",
                str_detect(
                    string = referral_source,
                    pattern = "northwest comm"
                ) ~ "Northwest Community Hospital, NorthShore",
                str_detect(
                    string = referral_source,
                    pattern = "revive"
                ) ~ "ReVive Center for Housing and Healing, Episcopal Charities",
                str_detect(
                    string = referral_source,
                    pattern = "west sub"
                ) ~ "West Suburban Medical Center, Pipeline Health",
                str_detect(
                    string = referral_source,
                    pattern = "cermak"
                ) ~ "Cermak Health Services of Cook County",
                str_detect(
                    string = referral_source,
                    pattern = "complex|ccdph"
                ) ~ "Cook County Department of Public Health",
            )
        ),
        primary_ref_type = as_factor(primary_ref_type),
        disposition = as_factor(disposition)
    ) 
```

Create summary table of referral sources:

```{r}
referral_sources <- referrals %>%
    group_by(referral_source_cat) %>%
    summarize(count = n()) %>%
    mutate(
        percent = round(
            (count/174*100),
            digits = 2
        ),
        variable = "referral_source"
    )
```

Create table of referral disposition totals:

```{r}
dispo_cat_totals <- referrals %>%
    group_by(
        disposition
    ) %>%
    summarize(
        dispo_cat_total = n()
    ) 
```

Create summary table of referral category counts grouped by disposition type:

```{r}
referral_types_by_dispo <- referrals %>%
    group_by(
        primary_ref_type,
        disposition
    ) %>%
    summarize(
        count = n()
    ) %>%
    full_join(
        dispo_cat_totals
    ) %>%
    mutate(
        percent = round(
            count/dispo_cat_total*100,
            digits = 2
        ),
        variable = "referral_type_by_dispo"
    ) %>%
    ungroup()
```

Create function for initial collection of encounter-level data:

```{r}
# Function takes a source file and desired columns as input arguments
collect_encounter_data <- function(source, columns) {
    # Print total number of records before processing
    print(
        paste0(
            "Initial number of records is ",
            nrow(source)
        )
    )
    # Filter for distinct records
    distinct1 <- distinct(source)
    # Print number of distinct records
    print(
        paste0(
            "Which reduces to ",
            nrow(distinct1),
            " when filtering for distinct rows"
        )
    )
    # Select only variables related to people-level data
    select2 <- select({{distinct1}}, {{columns}})
    # Filter for distinct people-level records
    distinct3 <- distinct(select2)
    # Print number of distinct people-level records
    print(
        paste0(
            "Which reduces to ",
            nrow(distinct3),
            " when filtering for distinct rows after selecting columns of interest"
        )
    )
    # Filter for only people actually admitted to the MRC
    filter4 <- filter(
        distinct3,
        # Criteria are their referral disposition was accepted and they have a last name in the record
        distinct3$disposition %in% "Accepted" & 
            !(is.na(distinct3$last_name))
    )
    # Print number of admitted clients
    print(
        paste0(
            "Which reduces to ",
            nrow(filter4),
            " when filtering for accepted clients"
        )
    ) 
    return(filter4)
}
```

Specify desired people-level variables:

```{r}
encounter_columns <- c(
    mrc_id = "Record ID", 
    last_name = "Client Last Name...55", 
    first_name = "Client First Name...56", 
    other_name = "Client Other name", 
    dob = "Client DOB...58", 
    sex = "Sex", 
    black = "What is your race? (choice=Black)", 
    white = "What is your race? (choice=White)",
    asian = "What is your race? (choice=Asian/American Indian or Alaska native)",
    pacific = "What is your race? (choice=Native Hawaiian or Pacific Islander)",
    multiracial = "What is your race? (choice=Multiracial)",
    other_race = "What is your race? (choice=Other)",
    hispanic = "Do you consider yourself Hispanic or Latinx?",
    mental_condition = "Do You have any mental health condition",
    hallucinations = "Do you hear voices or see things that others might not see?",
    hospitalized = "Have you ever been hospitalized for psychiatric or substance use conditions?",
    heroin_fentanyl = "heroin or fentanyl",
    pain_pills = "pain pills not prescribed to you",
    methadone = "methadone",
    bupe = "buprenorphine (Suboxone)",
    OUD_program = "Are you dispensed (methadone/burprenorphine) as part of a program?",
    alcohol_disorder = "Have you been hospitalized for dangerous alcohol withdrawal, had withdrawal seizures or hallucinations, or delirium tremens?",
    other_drugs = "Do you take any other street drugs?",
    referral_cat = "Primary Referral Category",
    care_need = "What is the care need?",
    insurance_cat = "What insurance do you have?",
    disposition = "Disposition?"
)
```

Run custom function defined above to collect admitted encounter records:

```{r}
mrc_just_encounters <- collect_encounter_data(source = all_raw, columns = encounter_columns)
```

Collect registration dates and join to the encounter table:

```{r}
mrc_encounters_admDate <- all_raw %>%
    select(
        mrc_id = "Record ID",
        mrc_admission_date = "MI1:Registration Date"
    ) %>%
    # Only use records associated with an encounter
    filter(mrc_id %in% mrc_just_encounters$mrc_id) %>%
    group_by(mrc_id) %>%
    # Filter out all the NA rows
    summarize(mrc_admission_date = first(mrc_admission_date)) %>%
    full_join(mrc_just_encounters)
```

Collect discharge dates and join to the encounter table:

```{r}
mrc_encounters_admDate_dischDate <- all_raw %>%
    select(
        mrc_id = "Record ID",
        mrc_discharge_date = "Discharge Date"
    ) %>%
    # Only use records associated with an admitted client and that aren't empty for a discharge date
    filter(
        mrc_id %in% mrc_just_encounters$mrc_id &
            !is.na(mrc_discharge_date)
    ) %>%
    group_by(mrc_id) %>%
    # Some clients had multiple records for one discharge, so merge those together
    summarize(mrc_discharge_date = max(mrc_discharge_date)) %>%
    full_join(mrc_encounters_admDate)
```

Collect discharge destinations and join to the encounter table:

```{r}
mrc_encounters <- all_raw %>%
    select(
        mrc_id = "Record ID",
        mrc_discharge_dest = "Discharge Destination"
    ) %>%
    # Only use records associated with an admitted client and that aren't empty for a discharge destination
    filter(
        mrc_id %in% mrc_just_encounters$mrc_id &
            !is.na(mrc_discharge_dest)
    ) %>% 
    group_by(mrc_id) %>%
    # Some clients had multiple records for one discharge, so merge those together
    summarize(mrc_discharge_dest = first(mrc_discharge_dest)) %>%
    full_join(mrc_encounters_admDate_dischDate) %>%
    mutate(
        # Coerce date variables into strings for easy interoperability with python
        across(
            .cols = c(dob, mrc_admission_date, mrc_discharge_date),
            .fns = ~ as.character(.x)
        )
    )
    
```

Some encounters needed manual chart review in order to fill in missing data on discharges:

```{r}
mrc_encounters[60,2] = "Unknown"
mrc_encounters[60,3] = "2021-03-27"
mrc_encounters[61,2] = "Rush Oak Park ED"
mrc_encounters[61,3] = "2021-03-12"
mrc_encounters[62,2] = "Unknown"
mrc_encounters[62,3] = "2021-03-09"
mrc_encounters[64,2] = "Housing Forward"
mrc_encounters[64,3] = "2021-04-30"
mrc_encounters[67,2] = "Unknown"
mrc_encounters[67,3] = "2021-05-23"
mrc_encounters[68,2] = "Stroger Hospital"
mrc_encounters[68,3] = "2021-07-03"
mrc_encounters[70,2] = "Unknown"
mrc_encounters[70,3] = "2021-09-09"
mrc_encounters[73,2] = "Unknown"
mrc_encounters[73,3] = "2021-08-22"
```

```{python}
print(r.mrc_encounters.head())
```

Pre-processing for summary stats:

```{r}
clustered_mrc_encounters <- clustered_mrc_encounters_raw %>%
    mutate(
        # Coerce date variables into the date object class
        across(
            .cols = c(dob, mrc_admission_date, mrc_discharge_date),
            .fns = ~ ymd(.x)
        ),
        # Capitalize name variables
        across(
            .cols = c(first_name, last_name, other_name),
            .fns = ~ str_to_title(.x)
        ),
        # Coerce these variables into un-ordered categories
        across(
            .cols = c(sex, referral_cat, insurance_cat, mrc_discharge_dest),
            .fns = ~ as_factor(.x)
        ),
        # Coerce these variables into booleans
        across(
            .cols = black:alcohol_disorder,
            .fns = ~ case_when(
                .x %in% c("checked", "yes") ~ TRUE,
                .x %in% c("unchecked", "no") ~ FALSE
            )
        ),
        # Calculate ages of clients at their admission dates
        age = floor(dob %--% mrc_admission_date / dyears(1)),
        # Convert ages into ordered age range categories 
        age_cat = factor(
            case_when(
                age < 25 ~ "<25",
                age > 24 & age < 35 ~ "25-34",
                age > 34 & age < 45 ~ "35-44",
                age > 44 & age < 55 ~ "45-54",
                age > 54 & age < 65 ~ "55-64",
                age > 64 ~ "65+"
            ),
            ordered = TRUE,
            levels = c("<25", "25-34", "35-44", "45-54", "55-64", "65+")
        ),
        # Synthesize a substance disorder variable
        substance_disorder = factor(
            case_when(
                heroin_fentanyl == TRUE |
                    pain_pills == TRUE |
                    methadone == TRUE |
                    bupe == TRUE |
                    OUD_program == TRUE |
                    alcohol_disorder == TRUE |
                    str_detect(
                        other_drugs,
                        "crack|marijuana|cocaine|heroin"
                    ) ~ "Yes",
                hospitalized == TRUE ~ "Maybe",
                TRUE ~ "No"
            ),
            ordered = TRUE,
            levels = c("No", "Maybe", "Yes")
        ),
        # Synthesize a mental disorder variable
        mental_disorder = case_when(
                mental_condition == TRUE |
                    hallucinations == TRUE |
                    hospitalized == TRUE |
                    substance_disorder > "No" ~ TRUE,
                TRUE ~ FALSE
        ),
        # Make all discharge destination strings lowercase for easier pattern matching
        mrc_discharge_dest = str_to_lower(mrc_discharge_dest),
        # Sythesize a discharge destination category variable
        discharge_dest_cat = as_factor(
            case_when(
                str_detect(
                    mrc_discharge_dest,
                    "bridge|sro|permanent|apartment|recovery|fillmore|transitional|greenwood|airport"
                ) ~ "Bridge/Stable Housing",
                str_detect(
                    mrc_discharge_dest,
                    "haymarket|rehab|(treatment center)|((?=.*h.a.s)(?=.*program))"
                ) ~ "Residential Treatment",
                str_detect(
                    mrc_discharge_dest,
                    "boulevard|julian"
                ) ~ "Interim/Emergency Housing, Chicago",
                str_detect(
                    mrc_discharge_dest,
                    "forward|((?=.*transfer)(?=.*room))|((?=.*shelter)(?=.*evanston))|rm|hf|((?=.*211)(?=.*fl))"
                ) ~ "Interim/Emergency Housing, Suburbs",
                str_detect(
                    mrc_discharge_dest,
                    "temporary|transfer|shelter"
                ) ~ "Interim/Emergency Housing, Unclear Region",
                str_detect(
                    mrc_discharge_dest,
                    "family|brother|sister|cousin|friend"
                ) ~ "Family/Friend/Other",
                str_detect(
                    mrc_discharge_dest,
                    "ltac|hospital|rush"
                ) ~ "ED/Hospital",
                str_detect(
                    mrc_discharge_dest,
                    "declined|abscond"
                ) ~ "Absconded/Without Plan",
                str_detect(
                    mrc_discharge_dest,
                    "home"
                ) ~ "Unclear",
                str_detect(
                    mrc_discharge_dest,
                    "unknown"
                ) ~ "Unknown"
            )
        )
    )
```

Make function to generate summary tables showing counts and percentages for different people-level variables:

```{r}
# Take source tibble and variable of interest as input
count_and_percent <- function(tbl, var_name) {
    tbl %>%
    group_by(.data[[var_name]]) %>%
    summarize(count = n()) %>%
    mutate(
        variable = var_name,
        percent = round(
            (count/nrow(tbl)*100),
            digits = 2
        )
    ) %>%
    rename(
        category = .data[[var_name]]
    )
}
```

Identify people-level variables for which you want summary tables:

```{r}
people_summary_names <- c("sex", "black", "white", "asian", "pacific", "multiracial", "other_race", "hispanic", "referral_cat", "insurance_cat", "age_cat", "mental_disorder", "substance_disorder", "discharge_dest_cat")
```

Automate the generation of a summary table for every single person-level variable of interest:

```{r}
people_summary_tables <- map(
    .x = people_summary_names, 
    .f = ~count_and_percent(clustered_mrc_encounters, .x)
)
```

Specify all the summary tables to be exported to a Google Sheet:

```{r}
all_summary_tables <- c(people_summary_tables, list(referral_sources, referral_cats_by_dispo))
all_summary_names <- c(people_summary_names, "referral_sources", "referral_cats_by_dispo")
```


Export data to Google Sheet:

```{r}
## Instantiate a Google Sheet if needed
# googlesheets4::gs4_create(
#     name = "mrc_summary_stats",
#     sheets = c(people_summary_names, "referral_sources", "referral_cats_by_dispo")
# )

purrr::map2(
    .x = all_summary_tables,
    .y = all_summary_names,
    .f = ~ googlesheets4::write_sheet(
        data = .x,
        ss = "https://docs.google.com/spreadsheets/d/1Set-aoj0jq4fCLemwu2tnj8rQsfOa8Ai-AIS8VK8it8/edit?usp=sharing",
        sheet = .y
    )
)
```

```{r}
referral_dests <- clustered_mrc_encounters %>% select(mrc_discharge_dest, discharge_dest_cat)
write_csv(referral_dests, "data/discharge_destinations.csv")
```

