---
title: "Compiling Summary Statistics for the Cook County Medical Respite Center"
author: "Daniel P. Hall Riggins, MD"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: html_document
---

# Load needed libraries:

```{r message=FALSE, warning=FALSE}
library(arrow)
library(lubridate)
library(tidyverse)
library(xlsx)
```

# Import data from RedCap:

```{r}
all_raw <- read_csv("data/MRC_ALL_DATA_LABELS_2021-09-30_1556.csv")
```
# MRC-Level Data

Calculate total months of operations:

```{r}
start_date <- as_date("2020-12-08") # From date of first communication
date_of_export <- as_date("2021-09-30") # From time-stamp on data file
operation_months <- start_date %--% date_of_export / dmonths(1)
print(
    paste0(
        "This data download represents ",
        floor(operation_months),
        " total months of operation."
    )
)
```

To get total referrals (174), get the max number for client ID:

```{r}
print(
    paste0(
        "There have been ",
        max(all_raw$`Record ID`),
        " total referrals."
    )
)
```
# Referral-Level Data

Collect referral-level data:

```{r}
referrals <- all_raw %>%
    select(
        record = `Record ID`, 
        referral_source = `Referring Organization`,
        primary_ref_type = `Primary Referral Category`,
        disposition = `Disposition?` # Aka whether referral was accepted
    ) %>%
    # Only keep records that have data present about one of the referral variables
    filter(
        !is.na(referral_source) | !is.na(primary_ref_type) | !is.na(disposition)
    ) %>%
    # Coerce variables into un-ordered factors 
    mutate(
        referral_source = str_to_lower(referral_source),
        # Consolidate referrals source categories using string matching
        referral_source_cat = as_factor(
            case_when(
                str_detect(
                    string = referral_source,
                    pattern = "macneal"
                ) ~ "MacNeal Hospital, Loyola Medicine",
                str_detect(
                    string = referral_source,
                    pattern = "loyola"
                ) ~ "Loyola UMC",
                str_detect(
                    string = referral_source,
                    pattern = "stroger"
                ) ~ "John H. Stroger Hospital, Cook County Health",
                str_detect(
                    string = referral_source,
                    pattern = "osf"
                ) ~ "OSF Little Company of Mary Medical Center",
                str_detect(
                    string = referral_source,
                    pattern = "olympia"
                ) ~ "Franciscan Health Olympia Fields",
                str_detect(
                    string = referral_source,
                    pattern = "alsip"
                ) ~ "CCDPH Alsip Alternate Housing Program",
                str_detect(
                    string = referral_source,
                    pattern = "ridge"
                ) ~ "Advocate Lutheran General Hospital",
                str_detect(
                    string = referral_source,
                    pattern = "swedish"
                ) ~ "Swedish Hospital, NorthShore",
                str_detect(
                    string = referral_source,
                    pattern = "gottlieb"
                ) ~ "Gottlieb Memorial Hospital, Loyola Medicine",
                str_detect(
                    string = referral_source,
                    pattern = "palos"
                ) ~ "Palos Health, Northwestern Medicine",
                str_detect(
                    string = referral_source,
                    pattern = "oak"
                ) ~ "Rush Oak Park Hospital",
                str_detect(
                    string = referral_source,
                    pattern = "connections"
                ) ~ "Connections for the Homeless",
                str_detect(
                    string = referral_source,
                    pattern = "provident"
                ) ~ "Provident Hospital of Cook County",
                str_detect(
                    string = referral_source,
                    pattern = "amita"
                ) ~ "AMITA Health Saint Francis Hospital Evanston",
                str_detect(
                    string = referral_source,
                    pattern = "rush u"
                ) ~ "Rush University Medical Center",
                str_detect(
                    string = referral_source,
                    pattern = "respond"
                ) ~ "Respond Now",
                str_detect(
                    string = referral_source,
                    pattern = "gateway"
                ) ~ "Gateway Foundation",
                str_detect(
                    string = referral_source,
                    pattern = "northwestern memorial"
                ) ~ "Northwestern Memorial Hospital",
                str_detect(
                    string = referral_source,
                    pattern = "safe"
                ) ~ "A Safe Haven",
                str_detect(
                    string = referral_source,
                    pattern = "defend"
                ) ~ "Law Office of the Cook COunty Public Defender",
                str_detect(
                    string = referral_source,
                    pattern = "access"
                ) ~ "Access Community Health Network",
                str_detect(
                    string = referral_source,
                    pattern = "sojou"
                ) ~ "Sojourner House",
                str_detect(
                    string = referral_source,
                    pattern = "northwest comm"
                ) ~ "Northwest Community Hospital, NorthShore",
                str_detect(
                    string = referral_source,
                    pattern = "revive"
                ) ~ "ReVive Center for Housing and Healing, Episcopal Charities",
                str_detect(
                    string = referral_source,
                    pattern = "west sub"
                ) ~ "West Suburban Medical Center, Pipeline Health",
                str_detect(
                    string = referral_source,
                    pattern = "cermak"
                ) ~ "Cermak Health Services of Cook County",
                str_detect(
                    string = referral_source,
                    pattern = "complex|ccdph"
                ) ~ "Cook County Department of Public Health",
            )
        ),
        primary_ref_type = as_factor(primary_ref_type),
        disposition = as_factor(disposition)
    ) 
```

# Referral-Level Summary Stats

Create summary table of referral sources:

```{r}
referral_sources <- referrals %>%
    group_by(referral_source_cat) %>%
    summarize(count = n()) %>%
    mutate(
        percent = round(
            (count/174*100),
            digits = 2
        ),
        variable = "referral_source"
    )
```

Create table of referral disposition totals:

```{r}
dispo_cat_totals <- referrals %>%
    group_by(
        disposition
    ) %>%
    summarize(
        dispo_cat_total = n()
    ) 
```

Create table of referral category counts grouped by disposition type:

```{r}
referral_types_by_dispo <- referrals %>%
    group_by(
        primary_ref_type,
        disposition
    ) %>%
    summarize(
        count = n()
    ) %>%
    full_join(
        dispo_cat_totals
    ) %>%
    mutate(
        percent = round(
            count/dispo_cat_total*100,
            digits = 2
        ),
        variable = "referral_type_by_dispo"
    ) %>%
    ungroup()
```
# Unclustered Encounter-Level Data

Create function for initial collection of encounter-level data:

```{r}
# Function takes a source file and desired columns as input arguments
collect_encounter_data <- function(source, columns) {
    # Print total number of records before processing
    print(
        paste0(
            "Initial number of records is ",
            nrow(source)
        )
    )
    # Filter for distinct records
    distinct1 <- distinct(source)
    # Print number of distinct records
    print(
        paste0(
            "Which reduces to ",
            nrow(distinct1),
            " when filtering for distinct rows"
        )
    )
    # Select only variables related to people-level data
    select2 <- select({{distinct1}}, {{columns}})
    # Filter for distinct people-level records
    distinct3 <- distinct(select2)
    # Print number of distinct people-level records
    print(
        paste0(
            "Which reduces to ",
            nrow(distinct3),
            " when filtering for distinct rows after selecting columns of interest"
        )
    )
    # Filter for only people actually admitted to the MRC
    filter4 <- filter(
        distinct3,
        # Criteria are their referral disposition was accepted and they have a last name in the record
        distinct3$disposition %in% "Accepted" & 
            !(is.na(distinct3$last_name))
    )
    # Print number of admitted clients
    print(
        paste0(
            "Which reduces to ",
            nrow(filter4),
            " when filtering for accepted clients"
        )
    ) 
    return(filter4)
}
```

Specify desired encounter-level variables:

```{r}
encounter_columns <- c(
    mrc_id = "Record ID", 
    last_name = "Client Last Name...55", 
    first_name = "Client First Name...56", 
    other_name = "Client Other name", 
    dob = "Client DOB...58", 
    sex = "Sex", 
    black = "What is your race? (choice=Black)", 
    white = "What is your race? (choice=White)",
    asian = "What is your race? (choice=Asian/American Indian or Alaska native)",
    pacific = "What is your race? (choice=Native Hawaiian or Pacific Islander)",
    multiracial = "What is your race? (choice=Multiracial)",
    other_race = "What is your race? (choice=Other)",
    hispanic = "Do you consider yourself Hispanic or Latinx?",
    mental_condition = "Do You have any mental health condition",
    hallucinations = "Do you hear voices or see things that others might not see?",
    hospitalized = "Have you ever been hospitalized for psychiatric or substance use conditions?",
    heroin_fentanyl = "heroin or fentanyl",
    pain_pills = "pain pills not prescribed to you",
    methadone = "methadone",
    bupe = "buprenorphine (Suboxone)",
    OUD_program = "Are you dispensed (methadone/burprenorphine) as part of a program?",
    alcohol_disorder = "Have you been hospitalized for dangerous alcohol withdrawal, had withdrawal seizures or hallucinations, or delirium tremens?",
    other_drugs = "Do you take any other street drugs?",
    referral_cat = "Primary Referral Category",
    care_need = "What is the care need?",
    insurance_cat = "What insurance do you have?",
    disposition = "Disposition?"
)
```

Run custom function defined above to collect admitted encounter records:

```{r}
just_encounters <- collect_encounter_data(source = all_raw, columns = encounter_columns)
```

Collect registration dates and join to the encounter table:

```{r}
encounters_admDate <- all_raw %>%
    select(
        mrc_id = "Record ID",
        mrc_admission_date = "MI1:Registration Date"
    ) %>%
    # Only use records associated with an encounter
    filter(mrc_id %in% just_encounters$mrc_id) %>%
    group_by(mrc_id) %>%
    # Filter out all the NA rows
    summarize(mrc_admission_date = first(mrc_admission_date)) %>%
    full_join(just_encounters)
```

Collect discharge dates and join to the encounter table:

```{r}
encounters_admDate_dischDate <- all_raw %>%
    select(
        mrc_id = "Record ID",
        mrc_discharge_date = "Discharge Date"
    ) %>%
    # Only use records associated with an admitted client and that aren't empty for a discharge date
    filter(
        mrc_id %in% just_encounters$mrc_id &
            !is.na(mrc_discharge_date)
    ) %>%
    group_by(mrc_id) %>%
    # Some clients had multiple records for one discharge, so merge those together
    summarize(mrc_discharge_date = max(mrc_discharge_date)) %>%
    full_join(encounters_admDate)
```

Collect discharge destinations and join to the encounter table:

```{r}
encounters <- all_raw %>%
    select(
        mrc_id = "Record ID",
        mrc_discharge_dest = "Discharge Destination"
    ) %>%
    # Only use records associated with an admitted client and that aren't empty for a discharge destination
    filter(
        mrc_id %in% just_encounters$mrc_id &
            !is.na(mrc_discharge_dest)
    ) %>% 
    group_by(mrc_id) %>%
    # Some clients had multiple records for one discharge, so merge those together
    summarize(mrc_discharge_dest = first(mrc_discharge_dest)) %>%
    full_join(encounters_admDate_dischDate) %>%
    mutate(
        # Coerce date variables into strings for easy interoperability with python
        across(
            .cols = c(dob, mrc_admission_date, mrc_discharge_date),
            .fns = ~ as.character(.x)
        ),
        name = paste0(first_name, " ", last_name)
    )
    
```

Some encounters needed manual chart review in order to fill in missing data on discharges:

```{r}

```

```{r}
write_parquet(encounters, "data/mrc_people_model/unclustered_MRC_encounters.parquet")
```

# Clustered Encounter-Level Data

```{r}
clustered_encounters_raw <- read_parquet("data/mrc_people_model/clustered_MRC_encounters.parquet")
```

Pre-processing for encounter/people-level summary stats:

```{r}
clustered_encounters <- clustered_encounters_raw %>%
    rename(cluster_id = `cluster id`) %>%
    mutate(
        mrc_id = as.integer(mrc_id)
    ) %>%
    # Add missing discharge information gotten from manual chart review
    rows_update(
        tibble(
            mrc_id = c(50L, 63L, 68L, 97L, 112L, 115L, 127L, 136L),
            mrc_discharge_dest = c("unknown","rush oak park ed","unknown","housing forward","unknown",
                "stroger hospital","unknown","unknown"),
            mrc_discharge_date = c("2021-03-07","2021-03-12","2021-03-09","2021-04-30","2021-05-23",
                "2021-07-03","2021-09-09","2021-08-22")
        ),
        by = "mrc_id"
    ) %>%
    # Correct mistake in sex entry
    rows_update(
        tibble(mrc_id = 160, sex = "male"), by = "mrc_id"
    ) %>%
    mutate(
        # Coerce date variables into the date object class
        across(
            .cols = c(dob, mrc_admission_date, mrc_discharge_date),
            .fns = ~ ymd(.x)
        ),
        # Capitalize name variables
        across(
            .cols = c(first_name, last_name, other_name),
            .fns = ~ str_to_title(.x)
        ),
        # Coerce these variables into un-ordered categories
        across(
            .cols = c(sex, referral_cat, mrc_discharge_dest),
            .fns = ~ as_factor(.x)
        ),
        # Coerce these variables into booleans
        across(
            .cols = black:alcohol_disorder,
            .fns = ~ case_when(
                .x %in% c("checked", "yes") ~ TRUE,
                .x %in% c("unchecked", "no") ~ FALSE
            )
        ),
        # Consolidate insurance categories
        insurance_cat = as_factor(
            case_when(
                str_detect(
                    string = insurance_cat,
                    pattern = "countycare|county care"
                ) ~ "CountyCare Health Plan, Medicaid",
                str_detect(
                    string = insurance_cat,
                    pattern = "meridian"
                ) ~ "Meridian, Medicaid",
                str_detect(
                    string = insurance_cat,
                    pattern = "illinicare"
                ) ~ "Aetna Better Health of Illinois, Medicaid",
                str_detect(
                    string = insurance_cat,
                    pattern = "blue cross community|bcbs medicaid"
                ) ~ "Blue Cross Community Health Plans, Medicaid",
                str_detect(
                    string = insurance_cat,
                    pattern = "omaha"
                ) ~ "Mutual of Omaha, Medicare Supplement",
                str_detect(
                    string = insurance_cat,
                    pattern = "united"
                ) ~ "United Healthcare, Medicaid",
                str_detect(
                    string = insurance_cat,
                    pattern = "wellcare"
                ) ~ "WellCare Health Plans, Medicare Advantage",
                str_detect(
                    string = insurance_cat,
                    pattern = "molina"
                ) ~ "Molina Healthcare of Illinois, Medicaid",
                str_detect(
                    string = insurance_cat,
                    pattern = "aetna"
                ) ~ "Aetna, Unspecified",
                str_detect(
                    string = insurance_cat,
                    pattern = "medicare"
                ) ~ "Medicare, Unspecified",
                str_detect(
                    string = insurance_cat,
                    pattern = "medicaid"
                ) ~ "Medicaid, Unspecified",
                str_detect(
                    string = insurance_cat,
                    pattern = "blue cross|bcbs"
                ) ~ "Blue Cross Blue Shield, Unspecified",
                TRUE ~ "Other"
            )
        ),
        # Calculate ages of clients at their admission dates
        age = floor(dob %--% mrc_admission_date / dyears(1)),
        # Convert ages into ordered age range categories 
        age_cat = factor(
            case_when(
                age < 25 ~ "<25",
                age > 24 & age < 35 ~ "25-34",
                age > 34 & age < 45 ~ "35-44",
                age > 44 & age < 55 ~ "45-54",
                age > 54 & age < 65 ~ "55-64",
                age > 64 ~ "65+"
            ),
            ordered = TRUE,
            levels = c("<25", "25-34", "35-44", "45-54", "55-64", "65+")
        ),
        # Synthesize a substance disorder variable
        substance_disorder = factor(
            case_when(
                heroin_fentanyl == TRUE |
                    pain_pills == TRUE |
                    methadone == TRUE |
                    bupe == TRUE |
                    OUD_program == TRUE |
                    alcohol_disorder == TRUE |
                    str_detect(
                        other_drugs,
                        "crack|marijuana|cocaine|heroin"
                    ) ~ "Yes",
                hospitalized == TRUE ~ "Maybe",
                TRUE ~ "No"
            ),
            ordered = TRUE,
            levels = c("No", "Maybe", "Yes")
        ),
        # Synthesize a mental disorder variable
        mental_disorder = case_when(
                mental_condition == TRUE |
                    hallucinations == TRUE |
                    hospitalized == TRUE |
                    substance_disorder > "No" ~ TRUE,
                TRUE ~ FALSE
        ),
        # Make all discharge destination strings lowercase for easier pattern matching
        mrc_discharge_dest = str_to_lower(mrc_discharge_dest),
        # Sythesize a discharge destination category variable
        discharge_dest_cat = as_factor(
            case_when(
                str_detect(
                    mrc_discharge_dest,
                    "bridge|sro|permanent|apartment|recovery|fillmore|transitional|greenwood|airport"
                ) ~ "Bridge/Stable Housing",
                str_detect(
                    mrc_discharge_dest,
                    "haymarket|rehab|(treatment center)|((?=.*h.a.s)(?=.*program))"
                ) ~ "Residential Treatment",
                str_detect(
                    mrc_discharge_dest,
                    "boulevard|julian"
                ) ~ "Interim/Emergency Housing, Chicago",
                str_detect(
                    mrc_discharge_dest,
                    "forward|((?=.*transfer)(?=.*room))|((?=.*shelter)(?=.*evanston))|rm|hf|((?=.*211)(?=.*fl))"
                ) ~ "Interim/Emergency Housing, Suburbs",
                str_detect(
                    mrc_discharge_dest,
                    "temporary|transfer|shelter"
                ) ~ "Interim/Emergency Housing, Unclear Region",
                str_detect(
                    mrc_discharge_dest,
                    "family|brother|sister|cousin|friend"
                ) ~ "Family/Friend/Other",
                str_detect(
                    mrc_discharge_dest,
                    "ltac|hospital|rush"
                ) ~ "ED/Hospital",
                str_detect(
                    mrc_discharge_dest,
                    "declined|abscond"
                ) ~ "Absconded/Without Plan",
                str_detect(
                    mrc_discharge_dest,
                    "home"
                ) ~ "Unclear",
                str_detect(
                    mrc_discharge_dest,
                    "unknown"
                ) ~ "Unknown"
            )
        )
    )
```

Consolidate the clustered people records

```{r}
people <- clustered_encounters %>%
    group_by(`cluster id`) %>%
    summarise(
        across(
            .cols = c(sex:hispanic, age_cat:mental_disorder),
            .fns = ~ first(.x)
        )
    )
```


```{r}
write_parquet(people, "data/mrc_people.parquet")
```

# Person and Encounter-Level Summary Stats

Make function to generate summary tables showing counts and percentages for different people-level variables:

```{r}
# Take source tibble and variable of interest as input
count_and_percent <- function(tbl, var_name) {
    tbl %>%
    group_by(.data[[var_name]]) %>%
    summarize(count = n()) %>%
    mutate(
        variable = var_name,
        percent = round(
            (count/nrow(tbl)*100),
            digits = 2
        )
    ) %>%
    rename(
        category = .data[[var_name]]
    )
}
```

Identify encounter-level variables for which you want summary tables:

```{r}
encounter_summary_names <- c("referral_cat", "insurance_cat", "discharge_dest_cat", "age_cat")
```

Identify people-level variables for which you want summary tables:

```{r}
people_summary_names <- c("sex", "black", "white", "asian", "pacific", "multiracial", "other_race", "hispanic", "mental_disorder", "substance_disorder")
```

Automate the generation of a summary table for every single encounter or person-level variable of interest:

```{r}
encounter_summary_tables <- map(
    .x = encounter_summary_names, 
    .f = ~count_and_percent(clustered_encounters, .x)
)

people_summary_tables <- map(
    .x = people_summary_names, 
    .f = ~count_and_percent(people, .x)
)
```

Specify all the summary tables to be exported to a Google Sheet:

```{r}
all_summary_tables <- c(encounter_summary_tables, people_summary_tables, list(referral_sources, referral_types_by_dispo))
all_summary_names <- c(encounter_summary_names, people_summary_names, "referral_sources", "referral_types_by_dispo")
```

# Export

Export data to Google Sheet:

```{r}
## Instantiate a Google Sheet if needed
# googlesheets4::gs4_create(
#     name = "mrc_summary_stats",
#     sheets = c(people_summary_names, "referral_sources", "referral_cats_by_dispo")
# )

purrr::map2(
    .x = all_summary_tables,
    .y = all_summary_names,
    .f = ~ googlesheets4::write_sheet(
        data = .x,
        ss = "https://docs.google.com/spreadsheets/d/1Set-aoj0jq4fCLemwu2tnj8rQsfOa8Ai-AIS8VK8it8/edit?usp=sharing",
        sheet = .y
    )
)
```
Export raw data to Excel Doc:

```{r}
purrr::map2(
    .x = list(referrals, clustered_encounters, people),
    .y = c("referrals", "clustered_encounters", "people"),
    .f = ~ write.xlsx(
        x = .x,
        file = paste0("data/cleaned_mrc_data_",Sys.Date(),".xlsx"),
        sheetName = .y,
        append = TRUE
    )
)
```

